\documentclass[12pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{array}
\usepackage{enumitem}
\usepackage{parskip}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta}

\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt}
\setlist[itemize]{noitemsep,topsep=2pt,leftmargin=1.4em,label={--}}
\setlist[enumerate]{noitemsep,topsep=2pt,leftmargin=1.8em}

\begin{document}

\begin{center}
{\Large\textbf{IE 400 -- Term Project Report}}\\[4pt]
Spring 2026
\end{center}

\bigskip

%=======================================================================
\section*{\underline{Question 1: Museum Escape}}

\subsection*{Problem Definition}

Lara starts at cell $(1,1)$ and must reach the exit at cell $(8,8)$ on an $8\times8$ grid
in the minimum number of time steps. At each time step she may move one cell up, down,
left, or right, or stay in place. She cannot enter obstacle cells. Three security cameras
patrol the museum; if Lara occupies any camera-visible cell at the same time step, she is
caught.

\textbf{Obstacle cells:}
$(1,3),\,(2,2),\,(2,5),\,(3,7),\,(4,2),\,(4,3),\,(5,5),\,(6,4),\,(7,1),\,(7,4),\,(7,5)$

\subsection*{Camera Movement Tables}

\textbf{Camera 1}:

\begin{center}
\begin{tabular}{cccc}
\toprule
$t$ & Position & Direction & Visible Cells \\
\midrule
0   & $(3,3)$ & Right & $(3,3),(3,4),(3,5)$ \\
1   & $(3,4)$ & Right & $(3,4),(3,5),(3,6)$ \\
2   & $(3,5)$ & Left  & $(3,5),(3,4),(3,3)$ \\
3   & $(3,4)$ & Left  & $(3,4),(3,3),(3,2)$ \\
$4+$ & \multicolumn{3}{l}{repeats from $t=0$} \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Camera 2}:

\begin{center}
\begin{tabular}{cccc}
\toprule
$t$ & Position & Direction & Visible Cells \\
\midrule
0 & $(4,6)$ & Down  & $(4,6),(5,6)$ \\
1 & $(5,6)$ & Down  & $(5,6),(6,6)$ \\
2 & $(6,6)$ & Right & $(6,6),(6,7)$ \\
3 & $(6,7)$ & Left  & $(6,7),(6,6)$ \\
4 & $(6,6)$ & Up    & $(6,6),(5,6)$ \\
5 & $(5,6)$ & Up    & $(5,6),(4,6)$ \\
$6+$ & \multicolumn{3}{l}{repeats from $t=0$} \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Camera 3}:

\begin{center}
\begin{tabular}{cccc}
\toprule
$t$ & Position & Direction & Visible Cells \\
\midrule
0 & $(7,7)$ & NE & $(7,7),(6,8)$ \\
1 & $(6,8)$ & SW & $(6,8),(7,7),(8,6)$ \\
2 & $(7,7)$ & SW & $(7,7),(8,6)$ \\
3 & $(8,6)$ & SW & $(8,6),(7,7),(6,8)$ \\
$4+$ & \multicolumn{3}{l}{repeats from $t=0$} \\
\bottomrule
\end{tabular}
\end{center}

The combined camera-visible set at time $t$ is
\[
  F(t) = C_1[t \bmod 4]\;\cup\; C_2[t \bmod 6]\;\cup\; C_3[t \bmod 4].
\]

%-----------------------------------------------------------------------
\subsection*{Task 1: Integer Programming Formulation}

\textbf{Sets and Parameters}
\begin{itemize}
  \item $\mathcal{V}$: set of all non-obstacle cells $(r,c)$, where $r,c\in\{1,\ldots,8\}$
  \item $N(r,c)$: set of cells reachable in one step from $(r,c)$, including $(r,c)$ itself
        and its orthogonal neighbours in $\mathcal{V}$
  \item $F(t)$: set of cells visible by any camera at time $t$
  \item $T$: planning horizon (total number of steps); to be minimised
\end{itemize}

\textbf{Decision Variables}
\[
  x_{t,r,c} \in \{0,1\} \qquad t=0,1,\ldots,T,\quad (r,c)\in\mathcal{V}
\]
$x_{t,r,c}=1$ if Lara occupies cell $(r,c)$ at time step $t$, and $0$ otherwise.

\textbf{Objective Function}

The IP is solved as a \textbf{feasibility problem} for a fixed horizon $T^*$ that is
pre-computed by BFS over the time-expanded state space. The objective is therefore a
constant zero:
\[
  \min\; 0
\]
The model checks whether a feasible path of exactly $T^*$ steps exists; BFS guarantees
that no shorter feasible path can exist, so the first $T^*$ for which the IP is feasible
is optimal.

\textbf{Constraints}

\textbf{(C1) Initial position:}
\[
  x_{0,1,1}=1, \qquad x_{0,r,c}=0 \quad \forall\,(r,c)\in\mathcal{V},\;(r,c)\neq(1,1)
\]

\textbf{(C2) One cell per time step:}
\[
  \sum_{(r,c)\in\mathcal{V}} x_{t,r,c}=1 \qquad \forall\;t=0,1,\ldots,T
\]

\textbf{(C3) Valid movement:}
\[
  x_{t+1,r,c}\;\leq\!\!\sum_{(r',c')\in N(r,c)}\!\! x_{t,r',c'}
  \qquad \forall\;t=0,\ldots,T-1,\quad(r,c)\in\mathcal{V}
\]

\textbf{(C4) Camera avoidance:}
\[
  x_{t,r,c}=0 \qquad \forall\;t,\quad\forall\,(r,c)\in F(t),\quad(r,c)\neq(8,8)
\]
The exit cell $(8,8)$ is exempt: Lara may step there even if a camera sees it.

\textbf{(C5) Exit reached at horizon $T$:}
\[
  x_{T,8,8}=1
\]

\textbf{(C6) Binary domain:} $x_{t,r,c}\in\{0,1\}$

%-----------------------------------------------------------------------
\subsection*{Task 2: Base Case}

BFS over the time-expanded state space finds $T^*=14$. The IP is solved at $T=14$ with
Gurobi and all constraints (C1)--(C6) are verified to hold.

\medskip
\textbf{Result:} $T^*=14$ steps.

\begin{center}
\begin{tabular}{cc|cc|cc}
\toprule
$t$ & Cell & $t$ & Cell & $t$ & Cell \\
\midrule
 0 & $(1,1)$ &  5 & $(5,2)$ & 10 & $(8,4)$ \\
 1 & $(2,1)$ &  6 & $(6,2)$ & 11 & $(8,5)$ \\
 2 & $(3,1)$ &  7 & $(7,2)$ & 12 & $(8,6)$ \\
 3 & $(4,1)$ &  8 & $(8,2)$ & 13 & $(8,7)$ \\
 4 & $(5,1)$ &  9 & $(8,3)$ & 14 & $(8,8)$ \\
\bottomrule
\end{tabular}
\end{center}

Lara goes straight down column~1 to row~5, cuts across to column~2, descends to row~8,
then moves right to the exit. No camera is active along this path at the relevant time
steps.

%-----------------------------------------------------------------------
\subsection*{Task 3: Booby Trap}

\textbf{Additional rule:} if Lara visits cell $(5,1)$, she may never visit cell $(8,3)$
afterwards.

We introduce a cumulative binary variable $y_t\in\{0,1\}$, where $y_t=1$ means Lara has
visited $(5,1)$ at or before time $t$. The linearisation is:
\begin{align*}
  y_t &\;\geq\; x_{t,5,1} && \forall\;t \tag{trap triggered}\\
  y_t &\;\geq\; y_{t-1}   && \forall\;t\geq1 \tag{stays triggered}\\
  x_{t,8,3} &\;\leq\; 1-y_{t-1} && \forall\;t\geq1 \tag{block $(8,3)$ after trap}
\end{align*}

\textbf{Result:} $T^*=16$ steps. The base-case path visits $(5,1)$ at $t=4$, which would
then forbid $(8,3)$. The solver finds an alternative route avoiding $(8,3)$ at the cost of
2 extra steps.

\begin{center}
\begin{tabular}{cc|cc|cc}
\toprule
$t$ & Cell & $t$ & Cell & $t$ & Cell \\
\midrule
 0 & $(1,1)$            &  6 & $(5,3)$ & 12 & $(5,7)$ \\
 1 & $(2,1)$            &  7 & $(5,4)$ & 13 & $(5,8)$ \\
 2 & $(3,1)$            &  8 & $(4,4)$ & 14 & $(6,8)$ \\
 3 & $(4,1)$            &  9 & $(4,5)$ & 15 & $(7,8)$ \\
 4 & $(5,1)$ \textit{[trap]} & 10 & $(4,6)$ & 16 & $(8,8)$ \\
 5 & $(5,2)$            & 11 & $(4,7)$ &    & \\
\bottomrule
\end{tabular}
\end{center}

%-----------------------------------------------------------------------
\subsection*{Task 4: Exit Lock ($t<18$)}

\textbf{Additional rule:} the cells
$\mathrm{LOCKED}=\{(6,7),(7,6),(7,7),(8,6),(6,8)\}$
are forbidden before $t=18$:
\[
  x_{t,r,c}=0 \qquad \forall\;t<18,\quad(r,c)\in\mathrm{LOCKED}.
\]

\textbf{Result:} $T^*=20$ steps. Lara waits at $(1,1)$ for the first four time steps, then
follows a route through the interior, arriving at $(6,8)$ exactly at $t=18$ when the lock
is lifted.

\begin{center}
\begin{tabular}{cc|cc|cc}
\toprule
$t$ & Cell & $t$ & Cell & $t$ & Cell \\
\midrule
 0 & $(1,1)$ &  8 & $(5,3)$ & 16 & $(5,7)$ \\
 1 & $(2,1)$ &  9 & $(5,4)$ & 17 & $(5,8)$ \\
 2 & $(3,1)$ & 10 & $(4,4)$ & 18 & $(6,8)$ \\
 3 & $(4,1)$ & 11 & $(4,4)$ & 19 & $(7,8)$ \\
 4 & $(5,1)$ & 12 & $(4,5)$ & 20 & $(8,8)$ \\
 5 & $(5,2)$ & 13 & $(4,6)$ &    & \\
 6 & $(6,2)$ & 14 & $(4,7)$ &    & \\
 7 & $(5,2)$ & 15 & $(4,7)$ &    & \\
\bottomrule
\end{tabular}
\end{center}

%-----------------------------------------------------------------------
\subsection*{Task 5: Ghost Checkpoint ($t=3$)}

\textbf{Additional rule:} Lara must occupy cell $(4,4)$ at exactly $t=3$:
\[
  x_{3,4,4}=1.
\]

\textbf{Result: Infeasible.} BFS on the obstacle map (ignoring cameras) shows that the
minimum number of steps to reach $(4,4)$ from $(1,1)$ is~6. The Manhattan distance alone
is $|4-1|+|4-1|=6$, and the obstacle cells $(4,2)$ and $(4,3)$ block shorter detours,
making the actual minimum at least~6. Since the constraint requires arrival at $t=3<6$,
it is geometrically impossible regardless of the chosen path. The model has no feasible
solution.

\textbf{Extended Result (Teleport Allowed):} If the movement constraint for the
transition $t=2 \to t=3$ is relaxed (i.e.\ a ``ghost teleport'' is permitted),
the model becomes feasible and yields $T^* = 14$ steps. Lara waits at $(1,1)$
for two time steps, teleports to $(4,4)$ at $t=3$, then proceeds to the exit.

\begin{center}
\begin{tabular}{cc|cc|cc}
\toprule
$t$ & Cell & $t$ & Cell & $t$ & Cell \\
\midrule
 0 & $(1,1)$ &  5 & $(5,3)$ & 10 & $(8,5)$ \\
 1 & $(2,1)$ &  6 & $(6,3)$ & 11 & $(8,5)$ \\
 2 & $(1,1)$ &  7 & $(7,3)$ & 12 & $(8,6)$ \\
 3 & $(4,4)$ \textit{[teleport]} &  8 & $(8,3)$ & 13 & $(8,7)$ \\
 4 & $(5,4)$ &  9 & $(8,4)$ & 14 & $(8,8)$ \\
\bottomrule
\end{tabular}
\end{center}
%-----------------------------------------------------------------------
\subsection*{Task 6: Branch and Bound (Base Case)}

\textbf{Branching Strategy.}  
If the LP relaxation gives a fractional solution, we branch on a fractional variable
$x_{t,r,c}$ by creating two subproblems:
\[
x_{t,r,c} = 0 \quad \text{and} \quad x_{t,r,c} = 1.
\]

\textbf{Bounding Strategy.}  
At each node, we solve the LP relaxation.  
\begin{itemize}
    \item If the solution is integer $\rightarrow$ it is feasible for the IP and becomes the incumbent.
    \item If it cannot improve the current solution $\rightarrow$ the node is pruned.
\end{itemize}

\textbf{Result for This Problem.}  
Since $T^* = 14$ is fixed by BFS, the IP becomes a feasibility problem. The LP relaxation
at the root node yields a \emph{fractional} solution (e.g.\ $x_{2,2,1}=0.5$). The
algorithm therefore branches, following a Depth-First Search (DFS) strategy. After
exploring a total of \textbf{12 nodes}, an integer-feasible solution is found at
\textbf{depth~11}.

\medskip
\textbf{Search Tree (DFS, depth-first).}

\begin{center}
\begin{tikzpicture}[
  every node/.style={draw, rounded corners, align=center, font=\small, inner sep=5pt},
  level distance=14mm,
  sibling distance=52mm
]
\node {Root (depth 0)\\ LP fractional\\ Branch on $x_{2,2,1}$}
  child { node {$x_{2,2,1}=0$ (depth 1)\\ LP fractional\\ $\vdots$}
    child { node {$\cdots$ (depth 11)\\ Integer solution\\ $\Rightarrow$ Optimal} }
    child[missing] {}
  }
  child { node {$x_{2,2,1}=1$ (depth 1)\\ (not reached)} };
\end{tikzpicture}
\end{center}

\textbf{Conclusion.}  
The LP relaxation at the root node yielded a fractional solution. The algorithm utilized a
Depth-First Search (DFS) strategy and explored a total of \textbf{12 nodes} before finding
the optimal integer solution at \textbf{depth~11}. The resulting path is identical to the
BFS and IP solutions: $T^*=14$ steps.

%=======================================================================
\section*{\underline{Question 2: CCI Production Planning}}

\subsection*{Problem Definition}

Coca-Cola \.{I}\c{c}ecek (CCI) plans production at two plants -- Ankara~(A) and
Istanbul~(I) -- over $m=1,\ldots,6$ months. The objective is to minimize the total cost of
inventory holding, cross-region transportation, workforce wages, hiring, layoffs, and
bottling-mode switching, subject to demand, capacity, workforce, and cross-shipping
constraints.

\subsection*{Data}

\textbf{Monthly demand} $d_{j,m}$ (liters, L):

\begin{center}
\begin{tabular}{lcccccc}
\toprule
Region $j$ & $m=1$ & $m=2$ & $m=3$ & $m=4$ & $m=5$ & $m=6$ \\
\midrule
Western (W)             & $9{,}000{,}000$ & $9{,}000{,}000$ & $14{,}000{,}000$ & $12{,}000{,}000$ & $12{,}000{,}000$ & $13{,}000{,}000$ \\
Central \& Eastern (CE) & $7{,}000{,}000$ & $7{,}000{,}000$ & $8{,}000{,}000$ & $9{,}000{,}000$ & $8{,}000{,}000$ & $9{,}000{,}000$ \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Plant parameters} ($i\in\{A,I\}$):

\begin{center}
\begin{tabular}{lll}
\toprule
Parameter & Ankara ($i=A$) & Istanbul ($i=I$) \\
\midrule
Bottling rate, 2\,L & $6{,}000$ bot/hr $=12{,}000$ L/hr & $9{,}000$ bot/hr $=18{,}000$ L/hr \\
Bottling rate, 1\,L & $10{,}000$ bot/hr $=10{,}000$ L/hr & $15{,}000$ bot/hr $=15{,}000$ L/hr \\
Available hours/month & 720 & 720 \\
Initial workforce $w_{i,0}$ & 60 & 100 \\
Primary region & CE & W \\
\bottomrule
\end{tabular}
\end{center}

\textbf{Cost parameters:}
\begin{itemize}
  \item Wage: $c_w=1{,}200$ TL/employee/month
  \item Hiring: $c_h=1{,}500$ TL/new employee
  \item Layoff: $c_l=2{,}000$ TL/laid-off employee
  \item Inventory: $c_{\mathrm{inv}}=0.05$ TL/L/month
  \item Cross-region transport: $c_{\mathrm{tr}}=0.10$ TL/L
  \item Switching: $c_{\mathrm{sw}}=20{,}000$ TL/plant/month
\end{itemize}

Policy parameters: $\alpha=0.25$, $\beta=0.10$, $\Delta=10$.

\textbf{Note.} The PDF states that 60\% of demand is in 2\,L bottles and 40\% in 1\,L
bottles. Since the problem adds ``\textit{you may split accordingly},'' we treat total
demand as a single aggregate and let the optimizer choose the 1\,L/2\,L mix freely; this
is already reflected in the bottling-rate constraint (D1).

\subsection*{Decision Variables}

Indices: $i\in\{A,I\}$ plant, $j\in\{W,CE\}$ region, $m\in\{1,\ldots,6\}$ month.
All production and inventory quantities in liters (L).

\begin{itemize}
  \item $p^1_{i,m},\,p^2_{i,m}\geq0$ \quad volume of 1\,L / 2\,L bottles produced at plant $i$ in month $m$ (L)
  \item $s_{i,j,m}\geq0$ \quad volume shipped from plant $i$ to region $j$ in month $m$ (L)
  \item $I_{i,m}\geq0$ \quad end-of-month inventory at plant $i$ in month $m$ (L)
  \item $w_{i,m}\in\mathbb{Z}_{\geq0}$ \quad packaging workforce at plant $i$ in month $m$
  \item $h_{i,m},\,l_{i,m}\in\mathbb{Z}_{\geq0}$ \quad employees hired / laid off
  \item $\delta_{i,k,m}\in\{0,1\}$ \quad 1 if bottle size $k\in\{1\text{L},2\text{L}\}$
        is produced at plant $i$ in month $m$
  \item $z_{i,m}\in\{0,1\}$ \quad 1 if both bottle sizes produced at plant $i$ in month $m$
\end{itemize}

\subsection*{Objective Function}

\[
\min\;Z = C_{\mathrm{inv}}+C_{\mathrm{tr}}+C_w+C_h+C_l+C_{\mathrm{sw}}
\]
\begin{align*}
C_{\mathrm{inv}} &= 0.05\sum_{i}\sum_{m}I_{i,m} \\
C_{\mathrm{tr}}  &= 0.10\!\!\!\sum_{\substack{i,j:\,j\neq\mathrm{primary}(i)}}\sum_{m}s_{i,j,m} \\
C_w  &= 1{,}200\sum_{i}\sum_{m}w_{i,m} \\
C_h  &= 1{,}500\sum_{i}\sum_{m}h_{i,m} \\
C_l  &= 2{,}000\sum_{i}\sum_{m}l_{i,m} \\
C_{\mathrm{sw}} &= 20{,}000\sum_{i}\sum_{m}z_{i,m}
\end{align*}

\subsection*{Constraints}

\textbf{(D1) Bottling capacity} (hours $\leq720$ per plant per month):
\[
  \frac{p^2_{A,m}}{12{,}000}+\frac{p^1_{A,m}}{10{,}000}\leq720,
  \qquad
  \frac{p^2_{I,m}}{18{,}000}+\frac{p^1_{I,m}}{15{,}000}\leq720
  \qquad\forall\,m
\]

\textbf{(D2) Packaging capacity:}
\[
  p^1_{i,m}+p^2_{i,m}\leq100{,}000\,w_{i,m} \qquad\forall\,i,m
\]

\textbf{(D3) Workforce balance} ($w_{A,0}=60$, $w_{I,0}=100$):
\[
  w_{i,m}=w_{i,m-1}+h_{i,m}-l_{i,m} \qquad\forall\,i,m
\]

\textbf{(D4) Workforce smoothing} ($\Delta=10$):
\[
  w_{i,m}-w_{i,m-1}\leq10, \qquad w_{i,m-1}-w_{i,m}\leq10 \qquad\forall\,i,m
\]

\textbf{(D5) Inventory balance} (initial inventories $=0$):
\[
  I_{i,m}=I_{i,m-1}+p^1_{i,m}+p^2_{i,m}-\sum_{j}s_{i,j,m} \qquad\forall\,i,m
\]

\textbf{(D6) Demand fulfillment} (no backlogging):
\[
  \sum_{i}s_{i,j,m}\geq d_{j,m} \qquad\forall\,j,m
\]

\textbf{(D7) Cross-shipping cap} ($\alpha=25\%$):
\[
  s_{i,j,m}\leq\alpha\,d_{j,m} \qquad\forall\,i,m,\;j\neq\mathrm{primary}(i)
\]

\textbf{(D8) Safety stock} ($\beta=10\%$, months $m=1,\ldots,5$):
\[
  I_{i,m}\geq\beta\,d_{j^*(i),\,m+1} \qquad m=1,\ldots,5 \quad\text{(quantities in L)}
\]
where $j^*(i)$ is the primary region of plant $i$.

\textbf{(D9) Zero ending inventory:}
\[
  I_{i,6}=0 \qquad\forall\,i
\]

\textbf{(D10) Switching cost} (big-$M=50{,}000{,}000$ L):
\[
  p^k_{i,m}\leq50{,}000{,}000\,\delta_{i,k,m} \qquad\forall\,i,k,m
\]
\[
  z_{i,m}\geq\delta_{i,1\mathrm{L},m}+\delta_{i,2\mathrm{L},m}-1 \qquad\forall\,i,m
\]

%-----------------------------------------------------------------------
\subsection*{Results}

The CCI model was solved to optimality with Gurobi. The optimal total minimum cost and its
breakdown by component are as follows:

\begin{center}
\begin{tabular}{lr}
\toprule
Cost Component & Amount (TL) \\
\midrule
Wage cost              & $1{,}414{,}800.00$ \\
Hiring cost            & $72{,}000.00$ \\
Layoff cost            & $12{,}000.00$ \\
Inventory holding      & $600{,}000.00$ \\
Cross-region transport & $70{,}000.00$ \\
Bottling switch cost   & $0.00$ \\
\midrule
\textbf{Total Minimum Cost} & $\mathbf{2{,}168{,}800.00}$ \\
\bottomrule
\end{tabular}
\end{center}

Notable findings from the optimal plan:
\begin{itemize}
  \item No bottling-mode switching cost is incurred ($z_{i,m}=0$ for all $i,m$); each plant
        uses at most one bottle size per month.
  \item Cross-region shipping is minimal (700{,}000\,L from Istanbul to
        Central~\&~Eastern in month~1 only), confirming that the 25\% cap is only
        occasionally binding.
  \item Workforce levels rise gradually from the initial 60 (Ankara) and 100
        (Istanbul) employees, respecting the $\Delta=10$ smoothing constraint
        throughout all six months.
\end{itemize}

Detailed monthly production, workforce, and inventory plans are generated by the Python
script and verified to satisfy all demands and constraints.

\end{document}
